---
title: "Species Dye Experiments"
output: pdf_document
date: "`r Sys.Date()`"
---

# 1.0 Load Data

## 1.1 Import Data
```{r}
source("Nectar_Dye_Data_Cleaning.R")

head(species_ND)
```

The main data file is called "species_ND", where ND stands for "Nectar Dye".


## 1.2 View Some Data Tables
```{r}
# View stigma by species
paste("Dye on stigma:")
table(species_ND$species, species_ND$dye_on_stigma)

# View reproductive structures by species
paste("Dye on reproductive structures:")
table(species_ND$species, species_ND$dye_on_rep_strucs)
```

# 2.0 Analyses

## 2.1 Foraging Strategy
```{r}
# Chi-Squared
chisq.test(table(species_ND$foraging, species_ND$dye_on_anthers_binary))
chisq.test(table(species_ND$foraging, species_ND$dye_on_rep_strucs))
chisq.test(table(species_ND$foraging, species_ND$dye_on_stigma_binary))

# Try with fishers exact test (small count in territorials with dye on stigma)
fisher.test(table(species_ND$foraging, species_ND$dye_on_anthers_binary))
fisher.test(table(species_ND$foraging, species_ND$dye_on_rep_strucs))
fisher.test(table(species_ND$foraging, species_ND$dye_on_stigma_binary))

# Cramer's V for effect size
library(rstatix)
cramer_v(table(species_ND$foraging, species_ND$dye_on_anthers_binary))
cramer_v(table(species_ND$foraging, species_ND$dye_on_stigma_binary))
cramer_v(table(species_ND$foraging, species_ND$dye_on_stigma_binary))
```
There is a significant difference in dye deposition between species based on foraging strategy in all cases.

## 2.2 Species
```{r}
# Dye on stigma based on species
chisq.test(table(species_ND$species, species_ND$dye_on_anthers_binary))
chisq.test(table(species_ND$species, species_ND$dye_on_rep_strucs))
chisq.test(table(species_ND$species, species_ND$dye_on_stigma_binary))

# Fisher's test
fisher.test(table(species_ND$species, species_ND$dye_on_anthers_binary))
fisher.test(table(species_ND$species, species_ND$dye_on_rep_strucs))
fisher.test(table(species_ND$species, species_ND$dye_on_stigma_binary))
```
There is a significant difference in dye deposition between species in all cases.

## 2.3 Sex
```{r}
# Chi-squared differences between sexes
chisq.test(table(species_ND$sex, species_ND$dye_on_anthers_binary))
chisq.test(table(species_ND$sex, species_ND$dye_on_rep_strucs))
chisq.test(table(species_ND$sex, species_ND$dye_on_stigma_binary))

# Fisher's test
fisher.test(table(species_ND$sex, species_ND$dye_on_anthers_binary))
fisher.test(table(species_ND$sex, species_ND$dye_on_rep_strucs))
fisher.test(table(species_ND$sex, species_ND$dye_on_stigma_binary))
```
There are no differences in dye deposition between sexes.