---
title: "Simulated Visits 2024 Stats"
output: pdf_document
date: "`r Sys.Date()`"
---

# 1.0 Load and Prep Data

## 1.1 Load Data

```{r message=FALSE}
source("Fruit_Set_Data_Cleaning_2024.R")
```

PRM_Trio_Strict: Subset of the data that contains pollinations done in March - June 2024 with 3 treatments:

- ND: Nectar Drop, renamed to Simulated Visits, contains only treatments with far pollen (not self or near-neighbour)
- HPO: Hand Pollination Only, renamed to hand pollination
- NE: Nectar Extraction

## 1.2 Prep Data

```{r}
# Add a new column to convert Fruit_Set from categorical (Yes, No) to binary (0,1)
PRM_Trio_Strict <- PRM_Trio_Strict %>% 
  mutate(Binary_FS = if_else(Fruit_Set=="Yes", 1, 0))
```

# 2.0 Run Model

## 2.1 GLMM

### Model Notes

- **Model:** Generalized Linear Mixed Model (GLMM) with a binomial error distribution and a logit link.
- **Response Variable:** Binary_FS (binary fruit set). Indicates whether a flower developed into a mature fruit (yes or no).
- **Explanatory Variable:** Treatment. There were 3 treatments listed above.
- **Random Effects:**
  - Plant_ID: Takes into account the inflorescence that the flower belonged to.
  - Patch: Takes into account inflorescence location.
- **Reference Group:** ND (nectar drop/simulated visits) was set as the reference group.

```{r}
# Set ND as the reference group
PRM_Trio_Strict$Treatment <- PRM_Trio_Strict$Treatment %>% factor(c("ND", "NE", "HPO"), levels = c("ND", "NE", "HPO"))

# Run model
glmm_PRM_Trio_Strict <- glmer(Binary_FS ~ Treatment + (1|Plant_ID) + (1|Patch), data=PRM_Trio_Strict, family=binomial(link="logit"))
summary(glmm_PRM_Trio_Strict)

MuMIn::r.squaredGLMM(glmm_PRM_Trio_Strict)
car::Anova(glmm_PRM_Trio_Strict, type="III", test.statistic="Chisq")
```

## 2.2 Check Model Assumptions
```{r}
library(DHARMa)
sim_res <- simulateResiduals(fittedModel = glmm_PRM_Trio_Strict)
plot(sim_res)

testUniformity(sim_res) # uniformity
testDispersion(sim_res) # dispersion
testZeroInflation(sim_res) # zero inflation
testOutliers(sim_res) # outliers

# Check residuals against predictors
plotResiduals(sim_res, PRM_Trio_Strict$Treatment)
plotResiduals(sim_res, PRM_Trio_Strict$Patch)
```

## 2.3 Odds Ratio
```{r}
# Pairwise ratios between treatments
library(emmeans)

emm <- emmeans(glmm_PRM_Trio_Strict, ~ Treatment)
emm_odds <- pairs(emm, type = "response")  
  # exponentiates automatically 
  # (takes the estimate and puts it to the exponent to get the odds)
emm_odds_df <- as.data.frame(emm_odds)
emm_odds_CI <- as.data.frame(confint(emm_odds))

# Merge tables
odds_ratio_df <- merge(emm_odds_df, 
                       emm_odds_CI, 
                       by = c("contrast", "odds.ratio", "SE", "df")) %>% 
  select(-c("null", "df"))

odds_ratio_df
```

